---
title: "Measuring network similarity over time"
author: "Tyler R. Bonnell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Measuring network similarity over time}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


##Introduction

  When looking at a social network in time, it is often important to ask how stable or similar is the network? This vignette runs through the use of lagged network measures in netTS. We start by introducing the cosine similarity measure proposed by Newman (2010), then introduce how to add custom functions.
  
The vignette is organised as follows:

1. Graph level similarity
2. Node level similarity
3. Application to real data
4. Custom functions for lagged measures

###Libraries
```{r, warning=FALSE,message=FALSE}
#if netTS is not yet installed you can use the following code to install it from github:
#devtools::install_github("tbonne/netTS")

library(ggplot2)
library(netTS)
library(igraph)
library(plyr)
library(dplyr)
library(reshape2)
library(lubridate)
```


###1. Graph level similarity

Simulate a random graph and create a list of modified graphs
```{r, fig.width=7, fig.height=6}
#1. create a random graph
graph.original <- igraph::erdos.renyi.game(15,0.3)

#2. simulate change to the original graph
graph.list <- list(graph.original)
graph.temp <- graph.original
for(i in 1:6){
  
  #remove and add an edge
  graph.modified <- graph.temp %>% delete.edges(sample(E(graph.temp),1)) 
  graph.modified <- graph.modified %>% add.edges(edges=c(sample(V(graph.modified),2))) 
  graph.modified <- simplify(graph.modified)
  
  #record change
  graph.list[[length(graph.list)+1]] <- graph.modified
  graph.temp <- graph.modified
}


#3. Plot the modified graphs
par(mfrow=c(2,3))
for(i in 1:6){
  plot(graph.list[[i]], main=paste0("Network ",i))
}
```

Measure similarity over time
```{r, fig.width=7}
#1. vector to store the similarity values
similarity.from.start <- vector()

#2. calculate similarity for each graph
for(i in 1:6){
  similarity.from.start[length(similarity.from.start)+1] <- cosine_between_graphs(graph.list[[1]],graph.list[[i]])
similarity.between.events <- vector()
}

#3. plot the similarity over time
plot(similarity.from.start, col="red", main="Network similarity", ylab = "Cosine similarity")
```


###2. Node level similarity

Using the same simulated data from above we now look at how similarity at the node level changes in time.


Measure similarity over time for nodes
```{r, warning=FALSE, fig.width=7}
#1. measure similarity at the node level
similarity.from.start <- data.frame((cosine_between_nodes(graph.list[[1]],graph.list[[1]])))
names(similarity.from.start) <- colnames(cosine_between_nodes(graph.list[[1]],graph.list[[1]]))
for(i in 2:6){
  similarity.from.start <- rbind.fill( similarity.from.start,  as.data.frame(cosine_between_nodes(graph.list[[1]],graph.list[[i]]))  )
}

#2. take a look at the similarity measures as a table
similarity.from.start #Note: NaN values here indicate when a node has no edges in at least one graph being compared.

#3. plot the similarity of nodes over time
similarity.from.start$seq <- seq(1,6,by=1)
df.melt<-melt(similarity.from.start, id.vars="seq")
names(df.melt)[2] <- "node"
ggplot(df.melt, aes(y=value, x= seq, color=node)) + geom_line() + theme_classic()

```


###3. Application to a real dataset: vervet grooming

Change in the network over time: relative to the first network in time
```{r, warning=FALSE,message=FALSE, fig.width=7}

#1. extract cosine measures at the graph level: compared to first network
graph.cosine <- graphTS(data = groomEvents, windowsize = days(30), windowshift = days(10), lagged = TRUE,firstNet = TRUE ,measureFun= cosine_between_graphs)

#2. extract cosine measures at the node level: compared to first network
node.cosine <- nodeTS(data = groomEvents, windowsize = days(30), windowshift = days(10), lagged = TRUE, firstNet=TRUE, measureFun= cosine_between_nodes)

#3. convert the data from wide format to long format for plotting
df.node.cosine <- melt(node.cosine, id.vars = c("windowstart","windowend", "nEvents" ))
names(df.node.cosine)[names(df.node.cosine)=="variable"] <- "node"

#4. plot the cosine similarities
ggplot(df.node.cosine, aes(x=windowstart, y=value, color=node)) + geom_line() + geom_line(data=graph.cosine, aes(x=windowstart, y=measure), size=2, col="black") + labs(y="Cosine similarity", x = "Day") + theme_classic()
```


###4. User specified functions for lagged comparisons

Create a user specified lagged network function. Here we just use the difference in mean strength between two networks.
```{r}
#the function should have two networks as the inputs
my.lagged.measure <- function(net_t1, net_t2){

  #Meaure between two networks
  diff.strength <- mean(strength(net_t2))-mean(strength(net_t1))
  
  #return value
  return(diff.strength)
  
}
```

Use the custom lagged function with netTS
```{r, fig.width=7}
#1. extract strength difference between graphs: compared to first network
graph.str.diff <- graphTS(data = groomEvents, windowsize = days(30), windowshift = days(10), lagged = TRUE,firstNet = TRUE ,measureFun= my.lagged.measure)

#2. extract strength difference between graphs: compared to previous network
graph.str.diff.prev <- graphTS(data = groomEvents, windowsize = days(30), windowshift = days(10), lagged = TRUE,firstNet = FALSE ,measureFun= my.lagged.measure)

#3. plot the strength differences from the first network
ggplot(graph.str.diff, aes(x=windowstart, y=measure)) + geom_line(col="blue") + labs(y="Strength similarity", x = "Day") + theme_classic() + geom_hline(yintercept = 0, linetype="dashed")

#4. plot the strength differences from the previous networks
ggplot(graph.str.diff.prev, aes(x=windowstart, y=measure)) + geom_line(col="blue") + labs(y="Strength similarity", x = "Day") + theme_classic() + geom_hline(yintercept = 0, linetype="dashed")

```



