---
title: "Measuring network similarity over time"
author: "Tyler R. Bonnell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Measuring network similarity over time}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


##Introduction

  When looking at a social network in time, it is often important to ask how stable or similar is the network? This vignette runs through the use of the cosine similarity measure proposed by Newman (2010), using the netTS package.
  
  The vignette is organised as follows:
  * 1. Graph level similarity
  * 2. Node level similarity
  * 3. Application to real data

###Libraries
```{r, warning=FALSE,message=FALSE}
#devtools::install_github("tbonne/netTS")
library(ggplot2)
library(netTS)
library(igraph)
library(plyr)
library(dplyr)
library(reshape2)
```


###1. Graph level similarity

Simulate a random graph and create a list of modified graphs
```{r}
#create a random graph
graph.original <- igraph::erdos.renyi.game(15,0.3)

#simulate change to the original graph
graph.list <- list(graph.original)
graph.temp <- graph.original
for(i in 1:6){
  
  #remove and add an edge
  graph.modified <- graph.temp %>% delete.edges(sample(E(graph.temp),1)) 
  graph.modified <- graph.modified %>% add.edges(edges=c(sample(V(graph.modified),2))) 
  graph.modified <- simplify(graph.modified)
  
  #record change
  graph.list[[length(graph.list)+1]] <- graph.modified
  graph.temp <- graph.modified
}

par(mfrow=c(2,3))
for(i in 1:6){
  plot(graph.list[[i]], main=paste0("Network ",i))
}
```

Measure similarity over time
```{r}
similarity.from.start <- vector()

for(i in 1:6){
  similarity.from.start[length(similarity.from.start)+1] <- cosine_between_graphs(graph.list[[1]],graph.list[[i]])
similarity.between.events <- vector()
}

plot(similarity.from.start, col="red", main="Network similarity", ylab = "Cosine similarity")
```


###1. Node level similarity

Using the same simulated data from above we now look at how similarity at the node level changes in time.


Measure similarity over time for nodes
```{r, warning=FALSE}
#measure similarity at the node level
similarity.from.start <- data.frame((cosine_between_nodes(graph.list[[1]],graph.list[[1]])))
names(similarity.from.start) <- colnames(cosine_between_nodes(graph.list[[1]],graph.list[[1]]))
for(i in 2:6){
  similarity.from.start <- rbind.fill( similarity.from.start,  as.data.frame(cosine_between_nodes(graph.list[[1]],graph.list[[i]]))  )
}

#take a look at the similarity measures as a table
similarity.from.start #Note: NaN values here indicate when a node has no edges in at least one graph being compared.

#plot the similarity of nodes over time
similarity.from.start$seq <- seq(1,6,by=1)
df.melt<-melt(similarity.from.start, id.vars="seq")
names(df.melt)[2] <- "node"
ggplot(df.melt, aes(y=value, x= seq, color=node)) + geom_line() + theme_classic()

```


###1. Application to a real dataset: vervet proximity network

Change in the network over time: relative to the first network in time
```{r, warning=FALSE,message=FALSE}
library(lubridate)
library(plyr)

#extract cosine measures at the graph level
graph.cosine <- graphTS(data = groomEvents, windowsize = days(30), windowshift = days(10), lagged = TRUE,firstNet = TRUE ,measureFun= cosine_between_graphs)

#extract cosine measures at the node level
node.cosine <- nodeTS(data = groomEvents, windowsize = days(30), windowshift = days(10), lagged = TRUE, firstNet=TRUE, measureFun= cosine_between_nodes)

#plot the cosine similarities
df.node.cosine <- melt(node.cosine, id.vars = c("windowstart","windowend", "nEvents" ))
names(df.node.cosine)[names(df.node.cosine)=="variable"] <- "node"
ggplot(df.node.cosine, aes(x=windowstart, y=value, color=node)) + geom_line() + geom_line(data=graph.cosine, aes(x=windowstart, y=measure), size=2, col="black") + labs(y="Cosine similarity", x = "Day") + theme_classic()
```


